Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-05-29 articles/formalism/sections/introduction.tex
Working some on introduction

\section{Introduction}
The Cosmic Mircowave Background is 10 percent polarized, but the polarization contains independent cosmological information.

Standard technique is to convert Stokes parameters into scalar and pseudo-scalar modes, which are easier to compare to theory.

The pseudo-scalar $B$ mode is particularly important because it cannot be generated by primordial scalar pertubations

primordial tensor perturbations

lensing of $E$-modes

systematics checks

exotic phenemena like cosmic birefringence

The spin-0 $E$/$B$ modes relate to the spin-2 Stokes parameters via the spin-raising and -lowering operators ($\eth,\bar \eth$), which are second derivatives evaulated locally.  However, in practice we compute $E$/$B$ modes to a specified band limit, and this makes them non-local functions of the polarization field.  In other words, the $E$/$B$ modes at a point can get contributions from all over the sky.

@@

Zaldarriaga explored the spatial real-space kernels in the flat-sky approximation \citep{}.

In this work we follow the convection in which bar-ed variables correspond to those in real space, while the tilde-ed variables correspond to those in harmonic space \cite{Zaldarriaga2001a}.

>>>>

\section{Introduction}
During reionization, the Cosmic Mircowave Background undergoes Thomson scattering that leaves it with $\sim 5$ percent linear polarization.  The polarization signal contains information about the plasma velocity and provides cosmological constraints independent from the signal in temperature \citep{1997NewA....2..323H}.  The standard analysis technique converts the Stokes parameters of polarization into scalar ($E$) and pseudo-scalar ($B$) modes, which are easier to compare to theory.  The pseudo-scalar $B$ mode is particularly important because it cannot be generated by primordial scalar pertubations.  Thus any $B$-modes that appear must be from:
% Need citations for all...
(1) primordial tensor perturbations (gravitational waves), which reveal the energy scale of inflation
\citep{1997PhRvD..56..596H,1997PhRvL..78.2054S};
(2) lensing of $E$-modes, which helps to reveal the distribution of matter in the universe;
(3) foregrounds (especially Galactic synchrotron and dust emission)
\citep{2016A&A...586A.133P};
(4) unknown systematic problems in the data;
(5) unknown, exotic phenemena like cosmic birefringence or primordial magnetic fields
\citep{1996ApJ...469....1K,1999PhRvL..83.1506L,2004ApJ...616....1C,2014MNRAS.438.2508P}.

The formalism for converting the Stokes parameters to scalar quantities is well established \citep{1997PhRvD..55.7368K,1997PhRvD..55.1830Z}.  The reason we take a detailed look at it here is that with renewed focus on foreground contamination to the $B$-mode signal, we wanted to gain an intuition for what $E/B$ mode patterns arise from physical polarized structures in the galaxy.  We also wanted to gain insight into the problem of masking and the non-locality of the $E/B$ fields compared to the Stokes parameters.

The spin-0 scalar $E$/$B$ modes relate to the spin-2 complex Stokes parameters via the spin-raising and -lowering operators ($\eth,\bar \eth$), which are second derivatives evaulated locally.  However, in practice we compute $E$/$B$ modes to a specified band limit, and this makes them non-local functions of the polarization field.  In other words, the $E$/$B$ modes at a point can get contributions from all over the sky.

@@

Zaldarriaga explored the spatial real-space kernels in the flat-sky approximation \citep{Zaldarriaga2001a}.

### 2018-05-29 articles/formalism/sections/real_space_formalism.tex
Working some on introduction

The standard construction of $E/B$ fields depend on the spin-raising and -lowering operators, and is usually carried out in harmonic space.  The spin-raising operator ($\eth$), applied to a field of spin-s $_{s}g$, results in a fields with spin-$(s+1)$: $(\eth _{s}g)' = e^{-i(s+1)\psi}(\eth _{s}g)$  \cite{goldberg67}.  The complementary spin-lowering operator $(\bar{\eth})$  similarly results with $(\bar{\eth} _{s}g)' = e^{-i(s-1)\psi}(\bar{\eth} _{s}g)$.  The complex spin-0 scalar now arise from  the spin-2 fields ${_{\pm 2}X}$ as follows,
%
\begin{subequations}\label{eq:ebdef}
\beqry
\mathcal{E}(\hat{n}) + i \mathcal{B}(\hat{n}) &=& -\bar{\eth}^2 _{+ 2}\bar{X}(\hat{n}) \,,\label{eq:ebdef_lower}\\
\mathcal{E}(\hat{n}) - i \mathcal{B}(\hat{n}) &=& -{\eth}^2 _{-2}\bar{X}(\hat{n}) \,.
\eeqry
\end{subequations}
%
%which by the virtue of being scalars are independent of coordinate definitions. %The real part of this complex spin-0 scalar field corresponds to the E-mode while the imaginary part to the B-modes \cite{Kamionkowski1997}. 
The $\cal E/B$ fields are defined locally at point $\hat n$ in terms of the operators $\eth$ and $\bar \eth$.

@@

We now define the different polarization data vectors and their representation in real and harmonic space as follows\footnote{We adopt a convention in which real space quantities are denoted by bar-ed variable while those in harmonic space are denoted by tilde-ed variables.}:
%
\beqrys
\bar{S} &=& \bmat E \\ B  \emat_{2 \rm N_{\rm pix} \times 1};\qquad \bar{X} = \bmat _{+2}X \\ _{-2}X \emat_{2 \rm N_{\rm pix} \times 1};\qquad \bar{P} =\fqu_{\tiny {2 \rm N_{\rm pix} \times 1}} \,, \\
\tilde{S} &=& \bmat a^{E} \\ a^{B} \emat _{2 \rm N_{\rm alms} \times 1};\qquad \tilde{X} = \bmat _{+2} \tilde{X} \\ _{-2} \tilde{X} \emat_{2 \rm N_{\rm alms} \times 1} \,.
\eeqrys
%
The different symbols have the same meaning as that discussed in \sec{sec:pol-primer}, except that the subscript ${\ell m}$ for the spherical harmonic coefficients is suppressed for cleaner notation.

>>>>

The standard construction of $E/B$ fields depend on the spin-raising and -lowering operators, and is usually carried out in harmonic space.
%
The spin-raising operator ($\eth$), applied to a field of spin-s $_{s}g$, results in a fields with spin-$(s+1)$: $(\eth _{s}g)' = e^{-i(s+1)\psi}(\eth _{s}g)$  \cite{goldberg67}.  The complementary spin-lowering operator $(\bar{\eth})$  similarly results with $(\bar{\eth} _{s}g)' = e^{-i(s-1)\psi}(\bar{\eth} _{s}g)$.  The complex spin-0 scalar now arise from  the spin-2 fields ${_{\pm 2}X}$ as follows,
%
\begin{subequations}\label{eq:ebdef}
\beqry
\mathcal{E}(\hat{n}) + i \mathcal{B}(\hat{n}) &=& -\bar{\eth}^2 _{+ 2}\bar{X}(\hat{n}) \,,\label{eq:ebdef_lower}\\
\mathcal{E}(\hat{n}) - i \mathcal{B}(\hat{n}) &=& -{\eth}^2 _{-2}\bar{X}(\hat{n}) \,.
\eeqry
\end{subequations}
%
%which by the virtue of being scalars are independent of coordinate definitions. %The real part of this complex spin-0 scalar field corresponds to the E-mode while the imaginary part to the B-modes \cite{Kamionkowski1997}. 
The $\cal E/B$ fields are defined locally at point $\hat n$ in terms of the operators $\eth$ and $\bar \eth$.

@@

We now define the different polarization data vectors and their representation in real and harmonic space as follows\footnote{We adopt a convention in which real space quantities are denoted by bar-ed variable while those in harmonic space are denoted by tilde-ed variables, as in \citep{Zaldarriaga2001a}. 
}:
%
\beqrys
\bar{S} &=& \bmat E \\ B  \emat_{2 \rm N_{\rm pix} \times 1};\qquad \bar{X} = \bmat _{+2}X \\ _{-2}X \emat_{2 \rm N_{\rm pix} \times 1};\qquad \bar{P} =\fqu_{\tiny {2 \rm N_{\rm pix} \times 1}} \,, \\
\tilde{S} &=& \bmat a^{E} \\ a^{B} \emat _{2 \rm N_{\rm alms} \times 1};\qquad \tilde{X} = \bmat _{+2} \tilde{X} \\ _{-2} \tilde{X} \emat_{2 \rm N_{\rm alms} \times 1} \,.
\eeqrys
%
The different symbols have the same meaning as that discussed in \sec{sec:pol-primer}, except that the subscript ${\ell m}$ for the spherical harmonic coefficients is suppressed for cleaner notation.
